\documentclass[11pt,a4paper,halfparskip]{scrartcl}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage{url} 
\usepackage[T1]{fontenc}
\usepackage{ucs} 
\pagestyle{plain}

\title{\small{Documentation for}\\\huge JULIELAB UIMA Wrapper for \\OpenNLP Part-of-Speech (POS)
Tagger}

\author{\normalsize Ekaterina Buyko\\
  \normalsize  Jena University Language \& Information Engineering (JULIE) Lab\\
  \normalsize F\"urstengraben 30 \\
  \normalsize D-07743 Jena, Germany\\
  {\normalsize \tt \{buyko\}@coling-uni-jena.de} }


\date{}

\begin{document}
\maketitle

\section{Objective}



The \textsc{OpenNLP POS Tagger}\footnote{http://www.opennlp.org} provides part of speech tags for
tokens. \textsc{JULIELAB UIMA Wrapper for OpenNLP POS Tagger} is part of the JULIE NLP tool
suite\footnote{\url{http://www.julielab.de/}} which contains several NLP components (all UIMA
compliant) from sentence splitting to named
entity recognition and normalization as well as a comprehensive UIMA
type system.

% now you can put a longer description of the tool here


%\section{Changelog}
% uncomment when needed


\section{Requirements and Dependencies}

% mostly our tools will be based on java 1.5 and use UIMA
\textsc{JULIELAB UIMA Wrapper for OpenNLP POS Tagger} is written in Java 1.5 using Apache
UIMA version 2.1.0-incubation\footnote{\url{http://incubator.apache.org/uima/}}. It
was not tested with other UIMA versions.

% ref to our type system
The input and output of an AE takes place by annotation objects. The
classes corresponding to these objects are part of a \emph{JULIE UIMA
  Type System}\footnote{The \emph{JULIE UIMA type system} can be
  obtained from \url{http://www.julielab.de/}}.

% now other dependencies you tool might have (you might organize this
% a subsections as well)




\section{Using the AE -- Descriptor Configuration}
% carefully edit this section!

In UIMA, each component is configured by a descriptor in XML. In the
following we describe how the descriptor required by this AE can be
created with \emph{Component Descriptor Editor}, an Eclipse plugin
which is part of the UIMA SDK.

A descriptor contains information on different aspects. The following
subsection refers to each sub aspect of the descriptor which is, in
the Component Descriptor Editor, a separate \emph{tabbed page}. For an
indepth description of the respective configuration aspects or tabs,
please refer to the \emph{UIMA SKD User's
  Guide}\footnote{\url{http://incubator.apache.org/uima/}}, especially
chapter 12 on ``Component Descriptor Editor User's Guide''.

To define your descriptor go through each tabbed pages mentioned
here, make your respective entries (especially in page \emph{Parameter
Settings} you will be able to configure \textsc{OpenNLP POS Tagger} 
to your needs) and save the descriptor as \\ \url{PosTagAnnotator.xml}.

\paragraph{Overview}
This tab provides general informtion about the component. For the
OpenNLP Sentence Splitter you need to provide the information as
specified in Table
\ref{tab:overview}.
% adapt to your needs, remember to change values in tabular below!

\begin{table}[h!]
  \centering
  \begin{tabular}{|p{3.5cm}|p{4cm}|p{6cm}|}
    \hline
    Subsection & Key & Value \\
    \hline\hline
    Implementation Details & Implementation Language &  \textsc{Java} \\
    \cline{2-3}
    & Engine Type & Primitive\\
    \hline
    Runtime Information & updates the CAS & yes  \\
    \cline{2-3}
    & multiple deployment allowed & yes \\
    \cline{2-3}
    & outputs new CASes & no \\
    \cline{2-3}
    & Name of the Java class file & \url{de.julielab.jules.ae.opennlp.PosTagAnnotator}\\
    \hline
    Overall Identification Information & Name & \textsc{JULIELAB UIMA Wrapper for OpenNLP Tokenizer}
\\
    \cline{2-3}
    & Version & 2.0 \\
    \cline{2-3}
    & Vendor & julielab\\
    \cline{2-3}
    & Description & see above\\
    \hline
  \end{tabular}
  \caption{Overview/General Settings for AE.}
  \label{tab:overview}
\end{table}


\paragraph{Aggregate}
% for primitive AEs this does not have to be set
Not needed here, as this AE is a primitive.

\paragraph{Parameters}
\label{sss:parameters}
% adapt to your needs

See Table \ref{tab:parameters} for a specification of the
configuration parameters of this AE. Do not check ``Use Parameter
Groups'' in this tab.

\begin{table}[h!]
  \centering
  \begin{tabular}{|p{4cm}|p{2cm}|p{2cm}|p{2cm}|p{4cm}|}
    \hline 
    Parameter Name & Parameter Type & Mandatory & Multivalued & Description \\
   \hline \hline
      modelFile & String & yes &no &  path to the \textsc{OpenNLP POS Tagger}
model.\\
 tagset & String & no &no &  CAS types to annotate (see \emph{JULIE UIMA type system}) \\
 language & String & yes & no &  language (e.g. eng) (see \emph{JULIE UIMA type system}) \\
 useTagdict & Boolean & yes & no &  \url{true} if a tag dictionary should be used \\
 tagDict & String & yes & no &  path to a tag dictionary (if the parameter value 
of \textsc{useTagdict} is \url{true})\\
 caseSensitive & Boolean & no & no &  \url{true} if a tag dictionary is case senstive \\
    % add your parameters here
  \hline
  \end{tabular}
  \caption{Parameters of this AE.}
  \label{tab:parameters}
\end{table}


\paragraph{Parameter Settings}
\label{sss:param_settings}
% adapt to your needs

The specific parameter settings are filled in here. For each of the
parameters defined in \ref{sss:parameters}, add the respective values
here (has to be done at least for each parameter that is defined as
mandatory). See Table \ref{tab:param_settings} for the respective
parameter settings of this AE.

\begin{table}[h!]
  \centering
  \begin{tabular}{|p{4cm}|p{4cm}|p{7cm}|}
    \hline
    Parameter Name & Parameter Syntax & Example \\
    \hline\hline
    modelFile & model.bin.gz & resources/POSTaggerPennBio.bin.gz\\
    tagset & CASType & de.julielab.jules.types.PennBioIEPOSTag \\
    language & ISO 639-1/2 & en \\
    useTagdict & Boolean & yes \\
    tagDict & dir/tagdict & resources/tagdictPennBioIE\\
   caseSensitive & Boolean & no \\
  \hline
  \end{tabular}
  \caption{Parameter settings of this AE.}
  \label{tab:param_settings}
\end{table}

\paragraph{Type System}
\label{sss:type_system}
On this page, go to \emph{Imported Type} and add the \emph{julie-morpho-syntax-types.xml} type system. (Use ``Import by Location'').


\paragraph{Capabilities}
% adapt if needed
Nothing needs to be done here.


\paragraph{Index}
% adapt if needed
Nothing needs to be done here.

\paragraph{Resources}
% adapt if needed
Nothing needs to be done here.


\section{Copyright and License}
% leave unchanged
This software is Copyright (C) 2007 Jena University Language \& Information
Engineering Lab (Friedrich-Schiller University Jena, Germany), and is
licensed under the terms of the Common Public License, Version 1.0 or (at
your option) any subsequent version.

The license is approved by the Open Source Initiative, and is
available from their website at \url{http://www.opensource.org}.

\bibliographystyle{alpha}
\bibliography{/home/paper/biblio/paper.bib}


\end{document}
